\chapter{Limitations of the Work}

\vspace*{1cm}

Although the project provides a useful comparison of optimization algorithms for
fine-tuning a Small Language Model, several limitations must be considered when
interpreting the results. These limitations are related to the dataset size, the
benchmark duration, the hardware environment, and the practical constraints of
the tested optimization methods.

\section{Limited Dataset Size}

The dataset used in this project was intentionally reduced in size in order to
make the experiments lighter and easier to execute. The processed training set
contains 2000 examples, while the validation and test sets contain 500 examples
each.

\vspace*{0.5cm}

This reduced dataset is sufficient for experimentation and optimizer comparison,
but it does not represent the full scale of instruction-tuning. Using a larger
dataset could lead to more stable training and more reliable conclusions about
the final model quality.

\section{Short Benchmark Duration}

The experimental benchmark was executed for 10 training steps with a batch size
of 4. This setup makes it possible to compare the early behavior of the
optimizers, but it is not enough to fully train the model.

\vspace*{0.5cm}

As a result, the reported losses and efficiency metrics should be interpreted as
a short benchmark rather than a complete fine-tuning experiment. A longer
training run with more epochs would be necessary to confirm the final ranking of
the optimizers.

\section{Hardware Dependency}

The fine-tuning experiments depend strongly on the hardware used. The results
were obtained on a machine equipped with a performant GPU. Metrics such as
training time, throughput, latency, and VRAM usage can change significantly on a
different GPU or on a CPU.

\vspace*{0.5cm}

Therefore, the numerical results are specific to the execution environment. The
general trends are still useful, but exact values should not be considered
universal for all machines.

\section{Energy Measurement Limitation}

Energy consumption is an important metric for Edge AI applications. However,
accurate energy measurement requires access to reliable power telemetry. When
this information is not available, the energy consumption metric cannot be fully
evaluated.

\vspace*{0.5cm}

In this project, the main comparison focuses on loss, latency, throughput, memory
usage, and model size. Energy consumption remains an important direction for
future experiments, especially when testing deployment on real edge hardware.

\section{Practical Limits of Second-Order Methods}

L-BFGS and Newton-CG are included because they are important optimization methods
from a theoretical point of view. However, they are not always practical for
transformer fine-tuning. They require more computation and memory than classical
first-order methods.

\vspace*{0.5cm}

The benchmark confirms this limitation. L-BFGS and Newton-CG show high latency
and high VRAM usage, which makes them less suitable for efficient language model
adaptation. Their inclusion remains useful for academic comparison, but they are
not the most realistic choices for deployment-oriented fine-tuning.

\section{No Final Compression Stage}

The title of the project includes optimization and compression of Small Language
Models. In the current implementation, the main completed part is the
fine-tuning and optimizer comparison. Compression techniques such as quantization,
pruning, or knowledge distillation are identified as future work.

\vspace*{0.5cm}

The results therefore provide a baseline for compression rather than a full
compressed model comparison. Future experiments should apply compression methods
to the best fine-tuned models and evaluate the impact on size, latency, memory,
and quality.

\section{Conclusion}

These limitations do not invalidate the project results, but they define the
scope of the conclusions. The work provides a controlled comparison of optimizer
behavior on \texttt{google/flan-t5-small}, while future work should extend the
experiments with longer training, larger datasets, real energy measurements, and
actual compression techniques.
