\begin{thebibliography}{99}

\bibitem{flant5}
H. W. Chung, L. Hou, S. Longpre, B. Zoph, Y. Tay, W. Fedus, E. Li, X. Wang,
M. Dehghani, S. Brahma, A. Webson, S. S. Gu, Z. Dai, M. Suzgun, X. Chen,
A. Chowdhery, A. Castro-Ros, M. Pellat, K. Robinson, D. Valter, S. Narang,
G. Mishra, A. Yu, V. Zhao, Y. Huang, A. Dai, H. Yu, S. Petrov, E. H. Chi,
J. Dean, J. Devlin, A. Roberts, D. Zhou, Q. V. Le, and J. Wei,
\textit{Scaling Instruction-Finetuned Language Models},
arXiv preprint arXiv:2210.11416, 2022.

\bibitem{t5}
C. Raffel, N. Shazeer, A. Roberts, K. Lee, S. Narang, M. Matena,
Y. Zhou, W. Li, and P. J. Liu,
\textit{Exploring the Limits of Transfer Learning with a Unified Text-to-Text Transformer},
Journal of Machine Learning Research, vol. 21, no. 140, pp. 1--67, 2020.

\bibitem{adam}
D. P. Kingma and J. Ba,
\textit{Adam: A Method for Stochastic Optimization},
International Conference on Learning Representations, 2015.

\bibitem{adamw}
I. Loshchilov and F. Hutter,
\textit{Decoupled Weight Decay Regularization},
International Conference on Learning Representations, 2019.

\bibitem{lbfgs}
D. C. Liu and J. Nocedal,
\textit{On the Limited Memory BFGS Method for Large Scale Optimization},
Mathematical Programming, vol. 45, pp. 503--528, 1989.

\bibitem{sgd}
L. Bottou,
\textit{Large-Scale Machine Learning with Stochastic Gradient Descent},
Proceedings of COMPSTAT, pp. 177--186, 2010.

\bibitem{transformers}
T. Wolf, L. Debut, V. Sanh, J. Chaumond, C. Delangue, A. Moi,
P. Cistac, T. Rault, R. Louf, M. Funtowicz, J. Davison, S. Shleifer,
P. von Platen, C. Ma, Y. Jernite, J. Plu, C. Xu, T. Le Scao, S. Gugger,
M. Drame, Q. Lhoest, and A. Rush,
\textit{Transformers: State-of-the-Art Natural Language Processing},
Proceedings of the 2020 Conference on Empirical Methods in Natural Language
Processing: System Demonstrations, pp. 38--45, 2020.

\bibitem{edgeai}
X. Wang, Y. Han, V. C. M. Leung, D. Niyato, X. Yan, and X. Chen,
\textit{Convergence of Edge Computing and Deep Learning: A Comprehensive Survey},
IEEE Communications Surveys \& Tutorials, vol. 22, no. 2, pp. 869--904, 2020.

\bibitem{alpaca}
Yahma,
\textit{alpaca-cleaned Dataset},
Hugging Face Datasets.
Available at: \url{https://huggingface.co/datasets/yahma/alpaca-cleaned}

\bibitem{flant5model}
Google,
\textit{google/flan-t5-small Model Card},
Hugging Face Models.
Available at: \url{https://huggingface.co/google/flan-t5-small}

\end{thebibliography}
