\chapter{Discussion}

\vspace*{1cm}

The experimental results show that optimizer choice has a strong influence on
both fine-tuning quality and computational efficiency. This chapter discusses the
main observations obtained from the benchmark and interprets them in the context
of Edge AI deployment.

\section{Quality and Efficiency Tradeoff}

The most important observation is the tradeoff between model quality and resource
efficiency. Adam achieves the best training and validation losses, which means it
is the most effective optimizer for improving the model during the benchmark.
However, Adam is also slower than SGD-based methods and produces higher inference
latency in the experiment.

\vspace*{0.5cm}

This result shows that the best optimizer depends on the target objective. If the
priority is to obtain the lowest validation loss, Adam is the strongest choice.
If the priority is to reduce latency, memory usage, and computational cost, SGD
or SGD with momentum may be more appropriate.

\section{Behavior of SGD-Based Methods}

SGD gives good efficiency results. It has low training step latency, high
throughput, and lower VRAM usage compared with the more complex optimizers. This
makes it attractive for constrained environments where computation and memory are
limited.

\vspace*{0.5cm}

SGD with momentum improves the validation loss compared with plain SGD while
keeping a reasonable latency and memory cost. This suggests that momentum helps
the optimizer move more effectively in the parameter space without adding too
much overhead. For Edge AI, this makes SGD with momentum a strong compromise
between performance and efficiency.

\section{Behavior of Adam and AdamW}

Adam clearly provides the best optimization quality in the benchmark. Its
adaptive learning rate mechanism allows it to update parameters more effectively
than SGD-based methods during the limited number of steps. This explains the
strong decrease in both training loss and validation loss.

\vspace*{0.5cm}

However, Adam requires additional optimizer states and has a higher computational
cost. The benchmark shows that it is slower and has higher inference latency than
SGD-based methods. AdamW with clipping and scheduling gives a more balanced
result. It does not reach the same validation loss as Adam, but it keeps better
efficiency and may represent a useful compromise in future experiments.

\section{Limits of Second-Order Methods}

L-BFGS and Newton-CG are important from a theoretical optimization perspective,
but the results show that they are less practical for this fine-tuning setup.
L-BFGS uses much more VRAM than first-order methods and does not significantly
improve the validation loss. Newton-CG is even more expensive, with the highest
step latency and the lowest throughput.

\vspace*{0.5cm}

These observations confirm that second-order or quasi-Newton methods are not
always suitable for transformer fine-tuning. Their theoretical advantages can be
limited by the high computational cost of applying them to models with many
parameters.

\section{Edge AI Interpretation}

In an Edge AI context, the final decision cannot be based only on validation loss.
The model must also be efficient enough to run under hardware constraints.
Latency, memory usage, and throughput are therefore essential criteria.

\vspace*{0.5cm}

From this perspective, Adam is useful as a quality baseline, but it may not be the
best choice if deployment efficiency is the main goal. SGD and SGD with momentum
are more attractive for constrained environments because they require less memory
and provide faster execution. AdamW with clipping and scheduling can also be
considered as a balanced candidate.

\section{Implications for Compression}

The results also provide a baseline for future compression work. Compression
techniques such as quantization or pruning should ideally start from a model that
already has good performance and acceptable efficiency. If the fine-tuned model is
too slow or too memory-intensive before compression, deployment may remain
difficult even after optimization.

\vspace*{0.5cm}

Adam can be selected when the objective is to maximize quality before applying
compression. On the other hand, SGD with momentum or AdamW may be selected when
the goal is to start from a model that already has a better balance between
quality and resource usage.

\section{Overall Discussion}

Overall, the benchmark confirms that there is no single optimizer that is best
for every criterion. Adam gives the best validation performance, SGD gives strong
efficiency, SGD with momentum improves the SGD baseline, and AdamW gives a
reasonable balance. L-BFGS and Newton-CG are less suitable for this practical
fine-tuning scenario.

\vspace*{0.5cm}

Therefore, the best optimizer should be chosen according to the target use case.
For pure fine-tuning quality, Adam is preferred. For Edge AI efficiency, SGD with
momentum or AdamW with clipping and scheduling may be more appropriate. This
tradeoff is central to the optimization and compression of Small Language Models.
